How much variance in AUC-ROC is there between model runs?

### 2020-06-06 rep_nosmooth/variance.Rmd
add quad model to variance check

```{r packages}
# Don't cache this, otherwise the chunks behind run-models will not work
suppressMessages({
  library(dplyr)
  library(readr)
  library(randomForest)
  library(pROC)
  library(foreach)
  library(doFuture)
  library(tidyr)
  library(ggplot2)
})
```

@@

features <- c("gov_opp_low_level", "gov_reb_low_level", "opp_gov_low_level",
              "reb_gov_low_level", "gov_opp_nonviol_repression", "gov_reb_nonviol_repression",
              "gov_opp_accommodations", "gov_reb_accommodations", "reb_gov_demands",
              "opp_gov_demands")

@@

model_table <- crossing(
  horizon = rep(c("1 month", "6 months"), SAMPLES),
  sample  = rep(1:SAMPLES, each = 2)
)

@@

  horizon_i <- model_table[i, ][["horizon"]]
  sample_i  <- model_table[i, ][["sample"]]

@@

  # randomForest with the x, y instead of formulate interface will not drop
  # missing values, so we need to make sure the training data drops those
  # manually
  train_data   <- train_data %>%
    dplyr::select(one_of(c(dv_name, features))) %>%
    filter(complete.cases(.))

  fitted_model <- suppressWarnings(randomForest(
    x = train_data[, features],
    y = train_data[, dv_name],
    ntree    = hp_settings$ntree,
    maxnodes = hp_settings$maxnodes,
    sampsize = hp_settings$sampsize,
    replace = FALSE, do.trace = FALSE, importance = FALSE
  ))

  # Because the original rep code uses the formula interface, the predictions
  # it puts out cover all of the test data, with NA values where needed for
  # incomplete rows. Replicate that here.
  non_missing_test_data <- test_data %>%
    dplyr::select(one_of(c("year", "period", "country_iso3", features))) %>%
    dplyr::filter(complete.cases(.))
  preds <- non_missing_test_data %>%
    dplyr::select(year, period, country_iso3) %>%
    mutate(pred = predict(fitted_model, newdata = non_missing_test_data,
                          type = "response")) %>%
    # add the non-complete rows back in
    right_join(test_data %>%
                 dplyr::select(year, period, country_iso3, one_of(dv_name)),
               by = c("year", "period", "country_iso3")) %>%
    # the original rep scripts do this; doesn't make a difference for ROC
    # calculations
    mutate(pred2 = ifelse(is.na(!!sym(dv_name)), NA_real_, pred)) %>%
    arrange(country_iso3, year, period)

@@

res_tbl <- res_long %>%
  group_by(horizon, smoothed) %>%
  dplyr::summarize(N = n(),
                   mean = mean(auc_roc),
                   min  = min(auc_roc),
                   max  = max(auc_roc),
                   sd   = sqrt(mean((auc_roc-mean(auc_roc))^2)), # pop sd, not sample sd
                   .groups = "drop")
res_tbl %>%
  knitr::kable(digits = 3)
```

@@

```{r}
# I want to add horizontal lines at the boundaries for rounding AUC-ROC to 
# two digits. Since the plot below is faceted, need to use a separate data frame
# for this with facet variable
hlines <- bind_rows(
  tibble(smoothed = "Smoothed", y = c(.82, .83, .84, .85) + .005),
  tibble(smoothed = "Original", y = c(.77, .78) + .005)
)

res_long %>%
  mutate(smoothed = ifelse(smoothed, "Smoothed", "Original"),
         horizon  = factor(horizon), 
         smoothed = factor(smoothed, levels = c("Smoothed", "Original"))) %>%
  ggplot() +
  facet_wrap(~smoothed, scales = "free") +
  geom_violin(aes(x = horizon, y = auc_roc)) +
  geom_point(aes(x = horizon, y = auc_roc), alpha = .5) +
  geom_hline(data = hlines, aes(yintercept = y), linetype = 3) +
  theme_bw()
```

>>>>

```{r packages}
# Don't cache this, otherwise the chunks behind run-models will not work
suppressMessages({
  library(dplyr)
  library(readr)
  library(randomForest)
  library(pROC)
  library(foreach)
  library(doFuture)
  library(tidyr)
  library(ggplot2)
  library(jsonlite)
})
```

@@

features <- fromJSON("output/model-definitions/feature-specs.json")

@@

model_table <- crossing(
  model   = c("Escalation", "Quad"),
  horizon = c("1 month", "6 months"),
  sample  = c(1:SAMPLES)
)

@@

  horizon_i  <- model_table[i, ][["horizon"]]
  sample_i   <- model_table[i, ][["sample"]]
  features_i <- features[[model_table[i, ][["model"]]]]

@@

  # randomForest with the x, y instead of formulate interface will not drop
  # missing values, so we need to make sure the training data drops those
  # manually
  train_data   <- train_data %>%
    dplyr::select(one_of(c(dv_name, features_i))) %>%
    filter(complete.cases(.))

  fitted_model <- suppressWarnings(randomForest(
    x = train_data[, features_i],
    y = train_data[, dv_name],
    ntree    = hp_settings$ntree,
    maxnodes = hp_settings$maxnodes,
    sampsize = hp_settings$sampsize,
    replace = FALSE, do.trace = FALSE, importance = FALSE
  ))

  # Because the original rep code uses the formula interface, the predictions
  # it puts out cover all of the test data, with NA values where needed for
  # incomplete rows. Replicate that here.
  non_missing_test_data <- test_data %>%
    dplyr::select(one_of(c("year", "period", "country_iso3", features_i))) %>%
    dplyr::filter(complete.cases(.))
  preds <- non_missing_test_data %>%
    dplyr::select(year, period, country_iso3) %>%
    mutate(pred = predict(fitted_model, newdata = non_missing_test_data,
                          type = "response")) %>%
    # add the non-complete rows back in
    right_join(test_data %>%
                 dplyr::select(year, period, country_iso3, one_of(dv_name)),
               by = c("year", "period", "country_iso3")) %>%
    # the original rep scripts do this; doesn't make a difference for ROC
    # calculations
    mutate(pred2 = ifelse(is.na(!!sym(dv_name)), NA_real_, pred)) %>%
    arrange(country_iso3, year, period)

@@

res_tbl <- res_long %>%
  group_by(smoothed, horizon, model) %>%
  dplyr::summarize(N = n(),
                   mean = mean(auc_roc),
                   min  = min(auc_roc),
                   max  = max(auc_roc),
                   sd   = sqrt(mean((auc_roc-mean(auc_roc))^2)), # pop sd, not sample sd
                   .groups = "drop")
res_tbl %>%
  knitr::kable(digits = 3)
```

@@

```{r}
# I want to add horizontal lines at the boundaries for rounding AUC-ROC to 
# two digits. Since the plot below is faceted, need to use a separate data frame
# for this with facet variable
hlines <- bind_rows(
  tibble(smoothed = "Smoothed", horizon = "1 month", 
         y = c(.79, .80, .81, .82, .83, .84, .85) + .005),
  tibble(smoothed = "Original", horizon = "1 month", 
         y = c(.77, .78, .79, .8) + .005),
  tibble(smoothed = "Smoothed", horizon = "6 months", 
         y = c(.77, .78, .79, .80, .81, .82) + .005),
  tibble(smoothed = "Original", horizon = "6 months", 
         y = c(.76, .77, .78, .79) + .005)
)

res_long %>%
  mutate(smoothed = ifelse(smoothed, "Smoothed", "Original"),
         horizon  = factor(horizon), 
         smoothed = factor(smoothed, levels = c("Smoothed", "Original"))) %>%
  ggplot() +
  facet_wrap(horizon ~ smoothed, scales = "free", nrow = 2) +
  geom_violin(aes(x = model, y = auc_roc)) +
  geom_point(aes(x = model, y = auc_roc), alpha = .5) +
  geom_hline(data = hlines, aes(yintercept = y), linetype = 3) +
  theme_bw() +
  coord_flip()
```
